
\usepackage[pdftex]{graphicx}


\usepackage{exercise}
\usepackage{enumitem}
\usepackage{color}
\usepackage{booktabs}

\newcommand{\pathname}[2]{\textit{ROOT}\texttt{/}\texttt{#1}\texttt{/}\texttt{#2}}

\begin{document}
\begin{center}
{\LARGE\bf Summer School on Formal Techniques\\ [1.5 ex]
Boolean (Un)Satisfiability \\[2ex]
\showexamname{2}}\\ [2 ex]
June, 2022
\end{center}
\section*{Explanation}

These exercises are designed to provide a deeper understanding of the
operation of Boolean satisfiability (SAT) solvers, especially when
applied to unsatisfiable formulas.  A key requirement is that solver
be able to generate a proof of unsatisfiability in such cases.

The provided problems range in how much time and effort is required,
and whether any programming is involved.  Each problem has an
associated {\em level}, according to the following standard:
\begin{description}
\item[I:] Simple pencil-and-paper exercises designed to provide a
  concrete examples for the concepts presented.  Doing these will
  help you gain confidence in the concepts being presented
\item[II:] More challenging pencil-and-paper exercises, or algorithmic
  and experimental activities.  These may running solvers on some
  benchmarks.
\item[III:] Deeper explorations.  These may require devising new
  algorithms, writing code, and performing experiments that go beyond
  the core lecture material.
\end{description}

All file names are specified in this document are given as path names
of the form
\pathname{\textit{DIR}}{\textit{FILE}}
where
\textit{ROOT} indicates the root of the directory structure,
\textit{DIR} is either ``\texttt{files}'' or ``\texttt{generators},''
and \textit{FILE} is the file name.


\newpage

\section*{Using the Provided Programs}

Here is how to use the provided tools.  In the following, we assume
{\it FILE} is the common base name for a set of files having different extensions.


\subsection*{TBSAT}

\begin{itemize}
\item Running in direct mode without proof generation, and generating up to $K$ solutions
\item[]\pathname{tbuddy/src/tbsat}{tbsat} \texttt{-i} {\it FILE}\texttt{.cnf} \texttt{-m} $K$
\item Running in direct mode with proof generation
\item[]\pathname{tbuddy/src/tbsat}{tbsat} \texttt{-i} {\it FILE}\texttt{.cnf} \texttt{-o} {\it FILE}\texttt{.lrat}
\item Running in bucket mode with proof generation
\item[]\pathname{tbuddy/src/tbsat}{tbsat} \texttt{-b} \texttt{-i} {\it FILE}\texttt{.cnf} \texttt{-o} {\it FILE}\texttt{.lrat}
\item Running with a schedule file and proof generation
\item[]\pathname{tbuddy/src/tbsat}{tbsat} \texttt{-s} {\it FILE}\texttt{.schedule} \texttt{-i} {\it FILE}\texttt{.cnf} \texttt{-o} {\it FILE}\texttt{.lrat}
\end{itemize}

\subsection*{LRAT-CHECK}

\begin{itemize}
\item Checking proof
\item[]\pathname{drat-trim}{lrat-check} {\it FILE}\texttt{.cnf} {\it FILE}\texttt{.lrat}
\end{itemize}



\newpage


\end{document}


